---
title: "Tips & Tricks:  Including Symbols, Superscripts, Subscripts & Line Breaks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{symbols}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Symbols

Symbols can be added to the label column via unicode.  Here is a quick reference to commonly used symbols.


```{r setup, echo = FALSE}
library(huxtable)

opath <- "."

df <- tibble::tribble(
  ~Symbol, ~`Textual Description`,  ~Unicode,  
  "\u2190", "Left arrow", "\\u2190", 
  "\u2192", "Right arrow", "\\u2192", 
  "\u2264", "Less-than or equal to", "\\u2264", 
  "\u2265", "Greater-than or equal to", "\\u2265", 
  "\u2260", "Not equal to", "\\u2260", 
  "\u00b1", "Plus-minus sign", "\\u00b1", 
  "\u03b1", "Alpha", "\\u03b1", 
  "\u03b2", "Beta", "\\u03b2", 
  "\u03bc", "Mu", "\\u03bc", 
  "\u00ab", "Non-breaking space", "\\u00ab"
)
df

# quick_rtf(hux(df),
#           file = "./test.rtf")
# 
# 
```


Here is an example call to tidytlg::gentlg() that will add the symbols to the label column.


```{r eval = FALSE}

df <- tibble::tibble(
  label = c("\u2264", "\u2265"),
  col1 = c("100", "200")
)

tidytlg::gentlg(df,
                file = "demo")

```


# Superscripts and Subscripts

Superscripts and Subscripts can be added to the label column via unicode.

```{r eval = FALSE}

df <- tibble::tibble(
  label = c("This is a superscript a{\\super a}", 
            "This is a subscript b{\\sub b}"),
  col1 = c("100", "200")
)

tidytlg::gentlg(df,
                file = "demo")

```


Superscripts and Subscripts can be added to the footnotes via unicode as well.

```{r eval = FALSE}

df <- tibble::tibble(
  label = c("This is a superscript a{\\super a}", 
            "This is a subscript b{\\sub b}"),
  col1 = c("100", "200")
)

tidytlg::gentlg(df,
                file = "demo",
                footers = "This is a footnote superscript{\\super a}")

```

# Inline RTF Line Breaks

Sometimes you need add a line break into your RTF.  Inserting '\\\\\\n' into your string will add your line break for you.


```{r eval = FALSE}

df <- tibble::tibble(
  label = c("Bodysystem \\\n Preferred Term"),
  col1 = c("100")
)

tidytlg::gentlg(df,
                file = "demo")


```

If you need a line break followed by a tab, just add in '\\\\li180'.

```{r eval = FALSE}

df <- tibble::tibble(
  label = c("Bodysystem\\\n\\li180Preferred Term"),
  col1 = c("100")
)

tidytlg::gentlg(df,
                file = "demo")


```

